``Interestingness'' in Mathematical Theory Exploration

### 2018-08-20 PhDThesis/evaluation_existingcode.tex
Some existing code changes

\section{Evaluation}
\label{sec:evaluation}

We have applied our recurrent clustering algorithm to several scenarios, with
mixed results. A major difficulty in evaluating these clusters is that we have
no ``ground truth'', i.e. there is no objectively correct way to compare
expressions. Instead, we provide a qualitative overview of the more interesting
characteristics.

@@

We have also applied our recurrent clustering algorithm to a variety of the
most-downloaded packages from \textsc{Hackage} (as of 2015-10-30), including
\texttt{text} (as above), \texttt{pandoc}, \texttt{attoparsec},
\texttt{scientific}, \texttt{yesod-core} and \texttt{blaze-html}. Whilst we
expected functions with a similar purpose to appear together, such as the
various reader and writer functions of \hs{pandoc}, there were always a few
exceptions which became separate for reasons which are still unclear.

@@

Whilst recurrent clustering has produced results which merit further
investigation, the application to theory exploration has yet to be tested
empirically. This is due to \qspec{}'s use of \qcheck{}'s \hs{Arbitrary} type
class to generate random values for instantiating variables. Whilst we can
automatically define \qspec{} theories and invoke them with \hs{nix-eval}, not
all types have \hs{Arbitrary} instances; those without cannot be given any
variables in our signature, which severely limits the possible combinations
which can be explored. In many cases, no variables can be included at all,
leaving just equations involving constants. This has so far prevented us from
measuring the direct impact on \qspec{} performance, either directly by
exploring the sub-sets identified through recurrent clustering, or indirectly by
comparing the equations generated by a full brute-force search to our recurrent
clusters: those equations relating terms from different clusters would not be
discovered by our method. It is this ratio of equations found through brute
force to those found after narrowing-down by clusters which is one of our key
objectives to maximise at this stage; until we begin to pursue the
\emph{interestingness} of the properties.

>>>>

To see how our recurrent clustering implementation performs outside our own
benchmark we have also applied it to existing code, with mixed results. Unlike
our benchmark, these scenarios have no ground truth or measure of success to
compare against (there is no objectively correct way to compare expressions).
Instead, we provide a qualitative overview of the more interesting
characteristics noticed in these explorations.

@@

For a more realistic test we also applied recurrent clustering to a variety of
the most-downloaded packages from \textsc{Hackage} (as of 2015-10-30), including
\texttt{text} (as above), \texttt{pandoc}, \texttt{attoparsec},
\texttt{scientific}, \texttt{yesod-core} and \texttt{blaze-html}. Whilst we
expected functions with a similar purpose to appear together, such as the
various reader and writer functions of \hs{pandoc}, there were always a few
exceptions which became separate for reasons which are still unclear.

@@

It is currently infeasible to extend these clustering applications into full
theory exploration runs, due to the use of constraints like \qcheck{}'s
\hs{Arbitrary} type class (or LeanCheck's , in the case of Speculate) to
generate suitable values for instantiating variables. Whilst we can
automatically define \qspec{} theories and invoke them with \hs{nix-eval}, not
all types have \hs{Arbitrary} instances; those without cannot be given any
variables in our signature, which severely limits the possible combinations
which can be explored. In many cases no variables can be included at all,
leaving just equations involving constants. This has so far prevented us from
measuring the direct impact on \qspec{} performance on real code examples,
either directly by exploring the sub-sets identified through recurrent
clustering or indirectly by comparing the equations generated by a full
brute-force search to our recurrent clusters.
